Windowed least-squares model reduction for dynamical systems

### 2019-10-14 main.tex
minor changes to 3.3

\section{Evaluation of gradients in the Euler--Lagrange equations for \methodAcronym\ with \spatialAcronym\ trial subspaces}\label{appendix:vector_calc}
We now derive the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation with \spatialAcronym\ trial subspaces. Without loss of generality, we present the derivation for a single window $t \in [0,T]$ with a constant basis $\basisspace$ and weighting matrix $\stweightingMat$.  
To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:euler_lagrange_appendix} for the integrand,
\begin{align*}\label{eq:integrand_apx}
 \minintegrand & \vcentcolon
(\genstateyDiscreteArgnt{}, \genstateyDiscreteDotArgnt{} ,\timeDummy) \mapsto \frac{1}{2} \big[
\basisspace \genstateyDiscreteDotArgnt{} - \velocity(\basisspace \genstateyDiscreteArgnt{}
+ \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[
\basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} +
\stateIntercept,t) \big], \\ & 
\vcentcolon \RR{\romdim} \times \RR{\romdim} \times [0,T]
 \rightarrow \RR{} .  
\end{align*}
To evaluate the gradients, we first expand $\minintegrand$:
\begin{align*}
 \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  &= \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} + \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[ \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \big] \\ 
 &= \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{} \big]  - \frac{1}{2}\big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{}\big] - \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \\ &+ \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
\end{align*}
Since $\stweightingMat$ is symmetric,
\begin{equation}\label{eq:int_expand}
 \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  = \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
\end{equation}
For notational purposes, we write the above as,
$$  \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) =  \minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy),$$
where
\begin{align*}
&\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv  \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big], \\
&   \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv   -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big], \\
&  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv  \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] . 
\end{align*}
Constructing the Euler--Lagrange equations for this functional $\minintegrand$ requires evaluating the derivatives $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and $\frac{\partial \minintegrand}{\partial \genstateyDiscreteDot}$. We start by evaluating $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and go term by term.

>>>>

\section{Evaluation of gradients in the Euler--Lagrange equations for \methodAcronym\ with \spatialAcronym\ trial subspaces}\label{appendix:vector_calc}
We now derive the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation with \spatialAcronym\ trial subspaces. Without loss of generality, we present the derivation for a single window $t \in [0,T]$ with a constant basis $\basisspace$ and weighting matrix $\stweightingMat$.  
To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:euler_lagrange_appendix} for the integrand
\begin{align*}\label{eq:integrand_apx}
 \minintegrand & \vcentcolon
(\genstateyDiscreteArgnt{}, \genstateyDiscreteDotArgnt{} ,\timeDummy) \mapsto \frac{1}{2} \big[
\basisspace \genstateyDiscreteDotArgnt{} - \velocity(\basisspace \genstateyDiscreteArgnt{}
+ \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[
\basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} +
\stateIntercept,t) \big], \\ & 
\vcentcolon \RR{\romdim} \times \RR{\romdim} \times [0,T]
 \rightarrow \RR{} .  
\end{align*}
To evaluate the gradients, we first expand $\minintegrand$:
\begin{align*}
 \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  &= \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} + \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[ \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \big] \\ 
 &= \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{} \big]  - \frac{1}{2}\big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{}\big] - \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \\ &+ \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
\end{align*}
Since $\stweightingMat$ is symmetric,
\begin{equation}\label{eq:int_expand}
 \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  = \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
\end{equation}
For notational purposes, we write the above as,
$$  \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) =  \minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy),$$
where
\begin{align*}
&\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv  \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big], \\
&   \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv   -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big], \\
&  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv  \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] . 
\end{align*}
Constructing the Euler--Lagrange equations for this functional $\minintegrand$ requires evaluating the derivatives $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and $\frac{\partial \minintegrand}{\partial \genstateyDiscreteDot}$. We start by evaluating $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and go term by term.

### 2019-10-14 section3/space_time.tex
through updates of Sec 3
fixing punctuation
minor changes to 3.3

\subsection{\spaceTimeAcronym\ trial spaces}\label{sec:wls_spacetime}
%\KTC{Error in first equation. This is the approximate solution not the real
%one. Use parallel language to hwo this was done in the last section for
%S-reduction trial subspaces. Also make sure the basis is zero at the beginning
%of the interval and the affine transformation is taken to be the solution at
%the end of the last window. This will not be too trivial, as the solution on
%the previous slab enters the trial subspace definition.}

%We now briefly consider \spaceTimeAcronym\ trial subspaces. 
The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
the FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$
with $\approxstateArgnt{n} \in \stspaceSTArg{n}$, where
\begin{equation}\label{eq:st_sttrialspace}
 \stspaceSTArg{n} \defeq
 \Range{\stbasisArgnt{n}} + \stateInterceptSTArg{n}\otimes \onesFunctionArg{n} \subseteq \RR{\fomdim} \otimes \timeSpaceArg{n}.
\end{equation}
Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix \textit{function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation. 
For \spaceTimeAcronym\ trial subspaces, we take the affine transformation to be $\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for $n=2,\ldots,\nslabs$ and $\stateInterceptSTArg{1} = \stateFOMIC$; this definition of the trial subspace automatically satisfies the boundary conditions.
At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as,
 \begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},
\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window. 
\begin{comment}
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptSTArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
\KTC{Remove all commas before equations unless it makes sense gramatically.
Add the $n$ superscript to $x_{ref}$}
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

\end{comment}
Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm} implies that \methodAcronym\ with \spaceTimeAcronym\ trial subspaces sequentially computes solutions $\stgenstateArg{n}$, $n = 1,\ldots,\nslabs$ that satisfy
\begin{align}\label{eq:obj_gen_slab_spacetime}
\begin{split}
&\underset{\stgenstateyArg{n} \in \RR{\stdim}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{n} + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} ).% \\ 
%      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n}  =0
%  \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}(\timeEndArg{n-1})\stgenstateArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMIC)
% & n=1. \end{cases}
%%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%\genstateIC & n=1, \end{cases} 
\end{split}
\end{align}
%\begin{align}\label{eq:obj_gen_slab_spacetime}
%\begin{split}
%&\stgenstateArg{n} = \underset{\stgenstatey \in \RR{\stdim}}{\text{arg\,min}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n} ), \\ 
%      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n}  =
%  \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0)
% & n=1. \end{cases}
%%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%%\genstateIC & n=1, \end{cases} 
%\end{split}
%\end{align}

>>>>

\subsection{\spaceTimeAcronym\ trial spaces}\label{sec:wls_spacetime}
\KTC{HERE}
%\KTC{Error in first equation. This is the approximate solution not the real
%one. Use parallel language to hwo this was done in the last section for
%S-reduction trial subspaces. Also make sure the basis is zero at the beginning
%of the interval and the affine transformation is taken to be the solution at
%the end of the last window. This will not be too trivial, as the solution on
%the previous slab enters the trial subspace definition.}

%We now briefly consider \spaceTimeAcronym\ trial subspaces. 
The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
the FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$
with $\approxstateArgnt{n} \in \stspaceSTArg{n}$, where
\begin{equation}\label{eq:st_sttrialspace}
 \stspaceSTArg{n} \defeq
 \Range{\stbasisArgnt{n}} + \stateInterceptSTArg{n}\otimes \onesFunctionArg{n} \subseteq \RR{\fomdim} \otimes \timeSpaceArg{n}.
\end{equation}
Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix \textit{function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation. 
For \spaceTimeAcronym\ trial subspaces, we take the affine transformation to be $\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for $n=2,\ldots,\nslabs$ and $\stateInterceptSTArg{1} = \stateFOMIC$; this definition of the trial subspace automatically satisfies the boundary condition at $t=0$ and enforces continuity between windows.
At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as
 \begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},
\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window. 
\begin{comment}
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptSTArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
\KTC{Remove all commas before equations unless it makes sense gramatically.
Add the $n$ superscript to $x_{ref}$}
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

\end{comment}
Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm} implies that \methodAcronym\ with \spaceTimeAcronym\ trial subspaces sequentially computes solutions $\stgenstateArg{n}$, $n = 1,\ldots,\nslabs$ that satisfy
\begin{align}\label{eq:obj_gen_slab_spacetime}
\begin{split}
&\underset{\stgenstateyArg{} \in \RR{\stdimArg{n}}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{} + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} ).% \\ 
%      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n}  =0
%  \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}(\timeEndArg{n-1})\stgenstateArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMIC)
% & n=1. \end{cases}
%%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%\genstateIC & n=1, \end{cases} 
\end{split}
\end{align}
%\begin{align}\label{eq:obj_gen_slab_spacetime}
%\begin{split}
%&\stgenstateArg{n} = \underset{\stgenstatey \in \RR{\stdim}}{\text{arg\,min}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n} ), \\ 
%      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n}  =
%  \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0)
% & n=1. \end{cases}
%%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%%\genstateIC & n=1, \end{cases} 
%\end{split}
%\end{align}
